---
title: "The effect of laughter on Perceived Stress"
author: "Konstantinos Pagonis"
date: "2022-12-21"
output: 
  github_document:
  toc: true
bibliography: 'References.bib'
---

***

```{r, error=FALSE, warning=FALSE, include=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(knitr)
library(haven)
library(broom)
library(tinytex)
library(broom)
```

# Introduction
It has been believed for centuries, that laughter can be benefitial for one's physical and mental health. In the end of the 20th century, scientists started providing evidence reflecting upon such claims. The Norman Cousins case (1979), where he used laughter and positive mood techniques in order to improve his "detoriating" health. Today, there is an official study of laughter called "Gelotology", and supports the claim that laughter can reduce stress chemicals in the human body Chang et al, 2013).A study by Eaton and Bradley (2008) that investigated differences between genders in stress appraisal, and concluded that females perceived the material significantly more stessful than the male participants. However, the differences of both genders in humour appreciation are not clear, and most studys are biased towards sex stereotypes (Gallivan, 1999).Other studies, where participants were exposed in humourous material, showed reduced anxiety and depression levels(Houston et al., 2010), and even improved blood vessel function(Miller & Fry, 2009). Overall, increasing amount of research suggests that the use of laughter can greatly improve health in several disorders. 

- H1:There will be an effect of laughter on perceived stress
- H2: There will be an effect of gender on perceived stress.
- H3: There will be an interaction effect of laughter and gender on perceived stress.


# Methods

## Design
The present study is a quantitative between subjects quasi-experiment. The effects of laughter (IV) on both males and females in perceived levels of stress (DV) will be investigated.The effect of laughter was measured by splitting the subjects in an experimental condition where the subjects were shows humorous images and a control condition where the images where neutral.

## Participants
The Participants were University students and were recruited via opportunity sampling via email list. There was a total of 40 participants, of which 20 were male and 20 female. The mean age of the participants was 34.62. Males and Females where split in equal numbers in each condition.

## Materials
The aim of this study is to present humorous images to participants in order to trigger laughter. An equal amount of images of humorous context for the laughter condition and 10 neutral images for the neutral condition. It was ensured that the images were non-offensive to any individual, and were found via research engine online. Then, the images were presented to the participants via a Powerpoint presentation.
The Perceived Stress Scale (PSS) was used to measure the perceived stress of the participants (Cohen et al., 1983). The PSS is a 10 item questionnaire which asks the participants about their affect in the last month and how often they felt that way. Other information such as gender and age were also collected through the questionnaire. The questionnaire likert scale included answers from 0-4, 0 means "never" and 4 means "very often". The final PSS scoring scale is from 0-40, the higher the score the higher the preceived stress.

## Procedure
Participants were given a Concent form and an debriefing sheet. The participants were informed about the steps that they would have to take for their contribution in this study. Such as that they will be shown some images and later they would complete a questionnaire. Once they agreed, and gave their signed consent, they would be shown humorous or neutral images depending on the group they were assigned for. Then they were provided with the PSS questionnaire to complete it. 

## Results


```{r}
library(readxl)
PSS1 <- read_excel("~/Desktop/Data Science/PSS1.xlsx")
View(PSS1)

```


## PSS Mean Scores Between Each Condition


```{r}
ggplot(PSS1, aes(x = Condition, y = PSSFINAL, fill = sex)) +
  labs(
    title = "PSS Mean Scores in Both Condition",
    x = "Conditions",
    y = "PSS_Mean_Scores"
  ) +
geom_boxplot(aes(color = sex))
```
 As the boxplot displays, women scored significantly higher on the PSS scores than males in both laughter and neutral conditions. Women in the Neutral Condition reported the highest scores of the study and men in the laughter condition the lowest. Also, a significant interaction between laughter and gender on perceived stress is displayed.


```{r}
twoANOVA <- aov(PSSFINAL ~ sex + Condition, data = PSS1)
summary(twoANOVA)
```

The Two-Way ANOVA reports that there was a significant effect of laughter on PSS scores    F(1,37) = 46, p<.001. Also, a significant effect of sex on Perceived Stress Score was also discovered F(1,37) = 49,  p<.001.


```{r}
interaction <- aov(PSSFINAL ~ sex*Condition, data = PSS1)
summary(interaction)
```








```{r}
mean(PSS1$age, na.rm = TRUE) #(arithmetic) Mean
sd(PSS1$age, na.rm = TRUE) #standard deviation

summary(PSS1$age)

```

# Discussion

According to our results, all the three hypothesis were accepted. A significant effect of laughter on perceived stress was found (H1), also, gender effect on perceived stress was observed (H2), and lastly, an interaction effect between laughter and gender on perceived stress (H3) due to the significantly higher scores of females in both conditions in comparison with that of males. Also, the scores of females were significantly lower in the laughter condition in comparison to the neutral condition.The findings of this study allign with the finding of previous studies (Chang et al., 2013; Houston et al., 2010) with a main difference that those studies focused on the effect of intense laughter. Jones et al.(2016), also supported that there were differences in their results between males and females in stress from a sample of university students. A limitation of the study is that there was no pre and post the participants were exposed to the images to received more accurate representation of the effet of laughter on perceived stress. Also, according to Chang et al.,(2013), laughter is a psychosocial response to a stimulus which includes positive physiological and psychological reaction. To conclude, the outcomes of this study suggest that laughter can prove fruitful and an economical way to improve stress levels and well being.

# References

[@chang2013] [@jones2016][@houston1998] [@khan2013][@eaton2008][@gallivan1999][@miller2009]
